---
title: "Preliminary exploratory analysis: PMD Alerts + Commits"
author: "Bruno Crotman"
date: "02/11/2019"
output: 
    pdf_document:
        toc: true
        number_sections: true
  
bibliography: bib.bib

    

---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}

library(git2r)
library(tidyverse)
library(magrittr)
library(furrr)
library(uuid)
library(lubridate)
library(gridExtra)
library(extrafont)
library(patchwork)

loadfonts()

knitr::opts_chunk$set(echo = FALSE)
```

\section{Introduction}

This document shows how to collect PMD alerts related to each commit of a project and introduces an analysis about this information.

In the Section \ref{PMD}, this document explains briefly how PMD Source Code Analyzer works and which kind of alerts it generates. In the Section \ref{prelim} the data collection and a brief analysis using the project Twitter4J is shown. The section \ref{conclusion} discusses the next steps for the work.

\section{PMD Alerts}\label{PMD}

PMD is a source code analyzer that finds common flaws, like unused variables, empty catch blocks, unnecessary object creation and so forth [@pmd].

In Java, the programing language in which Twittter4J is written, there are 8 groups of rules that generate alerts: 

* Best Practices: enforce generally accepted best practices, e.g. the usage of foreach instead of for when applicable, avoidance of hard coded IP addresses, avoidance of the reassignment of loop variables.

* Code Style: enforce specific ccode style, e.g, the non static classes must declare at leas one constructor, control statement must have braces.

* Design: e.g. abstract classses must have methods, classes must not have too many fields or methods

* Documentation: e.g. the comments' size must be within certain limits

* Error prone: detect constructs that are either broken, extremely confusing or prone to runtime errors, e.g. assigment to operands, if statements whose conditionals are always true or always false.

PMD lets you choose which types of alerts to use and even create new rules. In this work a "quick start" ruleset, that comes as a basic configuration, is used.


\section{Preliminary analysis}\label{prelim} 

In this preliminary analysis, the GitHub repository related to Twitter4J project [@twitter4j] is cloned and each commited code is analyzed by PMD.


\subsection{Collecting Data}\label{collect}


A script collects data from every commit made to Twitter4J project.

Each commit is checked out and the PMD alerts are generated.

In the end, all the PMD alerts are stored. 


```{r , eval=FALSE}



unlink("repository/Twitter4J", recursive = TRUE, force = TRUE)
  
clone("https://github.com/Twitter4J/Twitter4J.git","repository/Twitter4J")
 
repository <- repository("repository/Twitter4j")

iteracao <- 0

check_out_and_pmd <- function(commit){

    iteracao <<- iteracao + 1
    print(iteracao)
    temp <- (paste0(UUIDgenerate(),".csv"))
    print(commit$sha)    
    checkout(commit, force = TRUE)    
    shell(paste0("pmd/bin/pmd.bat -d repository/Twitter4J -f csv -R rulesets/java/quickstart.xml -cache cache -reportfile ", temp), shell = "PowerShell")
    conteudo <- read_csv(temp) %>% 
        mutate(sha = commit$sha)
    file.remove(temp)
    conteudo

}


line_count <- function (commit){

    iteracao <<- iteracao + 1
    print(iteracao)
    print(commit$sha)    
    checkout(commit, force = TRUE)    

    result <- shell("git -C repository/Twitter4J diff --stat 4b825dc642cb6eb9a060e54bf8d69288fbee4904", intern = TRUE ) %>% 
        enframe(value = "word") %>% 
        separate(col = word, into = c("file","number"), sep = "\\|") %>% 
        filter(str_detect(file, ".java")) %>% 
        mutate(lines = str_extract(number, "[0-9]+")) %>% 
        filter(!is.na(lines)) %>% 
        mutate(lines = as.integer(lines)) %>% 
        summarise(lines = sum(lines)) %>% 
        mutate(sha = commit$sha)
        
    
    result
    
}


check_out_and_pmd_possibly <- possibly(check_out_and_pmd, otherwise = tibble(sha = "erro"))

commits <-  commits(repository) %>%
    map_df(as_tibble)

saveRDS(commits, file =  "commits.rds")

commits_alerts <-  commits(repository) %>%
    map_dfr(check_out_and_pmd_possibly)

saveRDS(commits_alerts, file =  "commits_alerts.rds")

line_count_possibly <- possibly(line_count, otherwise = tibble(sha = "erro"))

line_counts <-  commits(repository) %>%
    map_df(line_count_possibly) 

saveRDS(line_counts, file =  "line_counts.rds")

tags_hist_author <- tags(repository) %>% 
  map(function(x) x$author ) %>% 
  map(function(x) as.character(x$when) ) %>%
  enframe() %>% 
  filter(str_detect(value,"[2]"))
  
tags_hist_tagger <- tags(repository) %>% 
  map(function(x) x$tagger ) %>% 
  map(function(x) as.character(x$when) ) %>%
  enframe() %>% 
  filter(str_detect(value,"[2]"))

tags_hist <- bind_rows(tags_hist_author,tags_hist_tagger) %>% 
  rename(
    version = name,
    date = value
  ) %>% 
  separate(col = version, into = c("major", "minor", "patch"), sep = "\\.") %>% 
  mutate(
    date = ymd_hms(date)
  ) %>% 
  arrange(date) %>% 
  mutate(
    lag_major = lag(major),
    lag_minor = lag(minor)
  ) %>% 
  mutate(
    major_changed = if_else(lag_major != major, T, F),
    minor_changed = if_else(lag_minor != minor | major_changed , T, F)
  ) 



saveRDS(tags_hist, file =  "tags_hist.rds")


```


Data can be loaded from the files that serve as cache.


```{r, cache=TRUE}

commits_from_cache <- read_rds("commits.rds")

alerts_from_cache <- read_rds("commits_alerts.rds")

lines_from_cache <- read_rds("line_counts.rds")

tags_from_cache <- read_rds("tags_hist.rds")


commits_com_tag <- commits_from_cache %>% 
  arrange(when) %>% 
  mutate(
    next_commit = lead(when)
  ) %>% 
  crossing(tags_from_cache) %>% 
  filter(date >= when, date < next_commit  ) %>% 
  select(
    sha,
    major,
    minor,
    patch,
    major_changed,
    minor_changed
  ) %>% 
  mutate(
    version = paste(major, minor, patch, sep = ".")
  )




all_data <-  commits_from_cache %>%  
    left_join(commits_com_tag, by = c("sha")) %>% 
    inner_join(alerts_from_cache, by = c("sha")) %>% 
    inner_join(lines_from_cache, by = c("sha")) 
  



```




\subsection{Evolution of the code size}


After an instability, it seems that the project evolves in a steady pace from the commit 300 until aproximately commit 1450, until the release of version 3.0.0. After that, the number of lines of code ceases to increase, and so do the number of alerts and the number of alerts per line of code.

Figure \ref{fig:evolution} shows this evolution, in terms of PMD alerts per line of code, total PMD alerts and total lines of code. The smaller red dots are releases of patch versions. The bigger and labeled ones are minor version releases. Here we use the definition of major, minor and patch versions as in Semantic Versioning 2.0.0 [@semver].



```{r, fig.cap="\\label{fig:evolution}Project evolution by commit"}


alerts_total <- all_data %>% 
    group_by(sha, when) %>% 
    summarise(
      alerts_per_line = n()/mean(lines), 
      lines = mean(lines), 
      alerts = n() ,
      version = first(version),
      major_changed = first(major_changed),
      minor_changed = first(minor_changed)
    ) %>% 
    ungroup() %>% 
    arrange(when) %>% 
    mutate(commit = row_number())

alerts_evolving <-  alerts_total %>% 
    filter(between(commit,300, 1450))


p1 <- ggplot(alerts_total) +
    geom_area(aes(x = commit, y = alerts_per_line ), fill = "gray") +
    geom_text(
      data = alerts_total %>%  filter(minor_changed),
      aes(x = commit, y = alerts_per_line + 0.006, label = version ),
      color = "red",
      size = 2,
      family = "CMU Serif"
    ) +  
    geom_area(data = alerts_evolving, aes(x = commit, y = alerts_per_line ), fill = "darkred") +
    geom_point(
      data = alerts_total %>% filter(minor_changed),
      aes(x = commit, y = alerts_per_line ),
      color = "red",
      size = 2
    ) +
    geom_point(
      data = alerts_total %>% filter(!is.na(version)),
      aes(x = commit, y = alerts_per_line ),
      color = "red",
      size = 1
    ) +
    theme_minimal() +
    scale_x_continuous(breaks = seq(0,to = 2500, by = 200)) +
    ggtitle("Alerts per line x Commit") +
    labs(x = "", y = "Alerts per line") +
    theme(text =  element_text(size = 8, family = "CMU Serif"))


p2 <- ggplot(alerts_total) +
    geom_area(aes(x = commit, y = alerts ), fill = "gray") +
    geom_area(data = alerts_evolving, aes(x = commit, y = alerts ), fill = "darkred") +
    geom_text(
      data = alerts_total %>%  filter(minor_changed),
      aes(x = commit, y = alerts + 300, label = version ),
      size = 2,
      color = "red",
      family = "CMU Serif"
    ) +
    geom_point(
      data = alerts_total %>% filter(minor_changed),
      aes(x = commit, y = alerts ),
      color = "red",
      size = 2
    ) +
    geom_point(
      data = alerts_total %>% filter(!is.na(version)),
      aes(x = commit, y = alerts ),
      color = "red",
      size = 1
    ) +
    theme_minimal() +
    scale_x_continuous(breaks = seq(0,to = 2500, by = 200)) +
    ggtitle("Alerts x Commit") +
    labs(x = "", y = "Alerts") +
    theme(text =  element_text(size = 8, family = "CMU Serif"))


p3 <- ggplot(alerts_total) +
    geom_area(aes(x = commit, y = lines ), fill = "gray") +
    geom_area(data = alerts_evolving, aes(x = commit, y = lines ), fill = "darkred") +
    geom_text(
      data = alerts_total %>%  filter(minor_changed),
      aes(x = commit, y = lines + 13000, label = version ),
      size = 2,
      color = "red",
      family = "CMU Serif"
    ) +
    geom_point(
      data = alerts_total %>% filter(minor_changed),
      aes(x = commit, y = lines ),
      color = "red",
      size = 2
    ) +
    geom_point(
      data = alerts_total %>% filter(!is.na(version)),
      aes(x = commit, y = lines ),
      color = "red",
      size = 1
    ) +
    theme_minimal() +
    scale_x_continuous(breaks = seq(0,to = 2500, by = 200))+
    ggtitle("Lines x Commit") +
    labs(x = "Commit", y = "Lines") +
    theme(text =  element_text(size = 8, family = "CMU Serif"))

p1 + p2 + p3 + plot_layout(ncol = 1) + plot_annotation(title = "Evolution along the commits", theme = theme(text =  element_text(size = 10, family = "CMU Serif")) )


```


Figure \ref{fig:quarters} shows the number of commits made in each quarter represented in the columns. 

The darkred columns are related to commits between 300 and 1450. These are the same commits represented in red in the Figure \ref{fig:evolution}, the period in which the number of lines of code is increasing.

The ribbons in the backgorund represent the number of releases. The color scale for the number of releases per quarter goes from white (no release) to dark red (more releases).

It´s possible to notice that there is a concentration in the same period when the number of lines and alerts per line is increasing.



```{r,  fig.cap="\\label{fig:quarters}Project evolution by quarter"}

commits_per_quarter <- alerts_total %>% 
  mutate(quarter = as.Date(round_date(when, "quarter")) ) %>% 
  group_by(quarter) %>% 
  count(n = n())

commits_per_quarter_evolving <- alerts_evolving %>% 
  mutate(quarter = as.Date(round_date(when, "quarter") )) %>% 
  group_by(quarter) %>% 
  count(n = n())

releases_per_quarter <- alerts_total %>%
  filter(!is.na(minor_changed)) %>% 
  mutate(quarter = as.Date(round_date(when, "quarter")) ) %>% 
  count(quarter) %>% 
  right_join(commits_per_quarter, by = c("quarter"), suffix = c("",".y")) %>% 
  mutate(n = if_else(is.na(n),0L,n)) %>% 
  select(quarter, n)
  
minor_releases <-  alerts_total %>%
  filter(minor_changed) %>% 
  mutate(quarter = as.Date(round_date(when, "quarter")) ) %>% 
  left_join(commits_per_quarter, by = c("quarter"))

  


  

ggplot(commits_per_quarter) +
  geom_col(
    data = releases_per_quarter, 
    aes(x = quarter, y = max(commits_per_quarter$n) * 1.2, fill = n ), 
    alpha = 0.5 , 
    position = "identity", 
    show.legend = FALSE
  ) +
  geom_col(aes(x = quarter, y = n)) +
  geom_col(
    data = commits_per_quarter_evolving, 
    aes(x = quarter, y = n), 
    fill = "darkred"
  ) +
  geom_text(
    data = minor_releases,
    aes(
      x = quarter,
      y = n,
      label = version
    ),
    nudge_y = 20,
    angle = 90,
    size = 3,
    family = "CMU Serif"
  ) +
  scale_x_date(
    date_breaks = "year",
    date_labels = "%Y"
  ) +
  ggtitle("Commits per quarter") +
  theme_minimal() +
  theme(text = element_text(family = "CMU Serif")) +
  labs(y = "Number of commits", x = "Quarter") +
  scale_y_continuous(limits = c(0,max(commits_per_quarter$n * 1.2))) +
  scale_fill_gradient(low = "white", high = "darkred" )






```

\newpage

\subsection{Types of alert}\label{sec:parts}


It's possible to analyze the amount of each type of alert in the code and how they evolve along the commits.

Figure \ref{fig:types} shows the evolution of the types of alerts by commit and releases. The light vertical lines refer to the release of a patch. The labeled vertical lines correspond to the minor and major releases.  

Code Style alerts are most common, and Error Prone alerts are the second most common at the end of the period. There is a fast growth in Error Prone alerts just before the release of the version 2.2.0.


```{r fig.cap="\\label{fig:types}Evolution of the types of alerts"}


releases <- alerts_total %>% 
  filter(!is.na(minor_changed))
  
releases_minor <- alerts_total %>% 
  filter(minor_changed)
  

rule_set <- all_data %>% 
    rename(
        rule_set = `Rule set`
    ) %>% 
    select(sha, when, author, Package, File, Priority, rule_set, Rule, lines) %>% 
    inner_join(alerts_total, by = c("sha")) %>% 
    mutate(per_line = 1/lines.x) %>% 
    group_by(commit, rule_set) %>% 
    summarise(per_line = sum(per_line)) 


ggplot(rule_set) +
    geom_line(aes(color = rule_set, x = commit, y = per_line), size = 1.2) +
    geom_vline(data = releases,aes(xintercept = commit ), color = "darkred", alpha = 0.4) +
    geom_vline(data = releases_minor,aes(xintercept = commit ), color = "darkred", alpha = 0.8) +
    geom_text(
      data = releases_minor,
      aes(x = commit, y = 1.2* max(rule_set$per_line), label = version ), 
      color = "darkred",
      alpha = 0.8,
      family = "CMU Serif",
      size = 3,
      angle = 90,
      nudge_x = -25
    ) +
    theme_minimal() +
    theme(
      legend.position = "top",
      text = element_text(family = "CMU Serif")
    ) +
    labs(x = "Commit", y = "Alerts per line", color = "Rule set")



```


\subsection{Priorities of the Alerts }

PMD assigns priorities to the alerts. Priority 1 is the worst.

Figure \ref{fig:priority} show the alerts by priority.

There is a big increase just before the release of version 2.2.0. The origin of this increase must linked with the increase in the Error Prone alerts refered in Section \ref{sec:parts}.


```{r fig.cap="\\label{fig:priority}Evolution of the alerts by priority"}

priority <- all_data %>% 
    rename(
        rule_set = `Rule set`
    ) %>% 
    select(sha, when, author, Package, File, Priority, rule_set, Rule, lines) %>% 
    inner_join(alerts_total, by = c("sha")) %>% 
    mutate(per_line = 1/lines.x) %>% 
    group_by(commit, Priority) %>% 
    summarise(per_line = sum(per_line)) %>% 
    ungroup() %>% 
    mutate(Priority = as.factor(Priority)) 


ggplot(priority) +
    geom_line(aes(color = Priority, x = commit, y = per_line), size = 1.25) +
    geom_vline(data = releases,aes(xintercept = commit ), color = "darkred", alpha = 0.4) +
    geom_vline(data = releases_minor,aes(xintercept = commit ), color = "darkred", alpha = 0.8) +
    geom_text(
      data = releases_minor,
      aes(x = commit, y = 1.2* max(priority$per_line), label = version ), 
      color = "darkred",
      alpha = 0.8,
      family = "CMU Serif",
      size = 3,
      angle = 90,
      nudge_x = -25
    ) +
    theme_minimal() +
    theme(
      legend.position = "top",
      text = element_text(family = "CMU Serif")
    ) +
    labs(x = "Commit", y = "Alerts per line", color = "Priority") +
    scale_color_manual(values = c("1" = "darkred","2" = "chocolate","3" = "orange","4" = "green")) 
  



```


\section{Conclusions and next steps} \label{conclusion}

In the next weeks we will dive into this data. Now that we have the tools we need to grab the data linking alerts and commits of the projects, we can go through some analysis:

* Can we find developers that consistently commit with too much alerts per line? 

* Is "Alerts per line" a good proxy for kludges ?

* Does the behavior of the developers change when they are aproaching a release or when the frequency of the commits is higher?

* Does the team change its behavior when a member starts to kludge?

* Does a change in the team behavior degrade the quality of the code?

* All of these analysis can be made for many java open source projects. Some of the projects we can analyse are: Elastic Search (https://github.com/elastic/elasticsearch), Hadoop (https://github.com/apache/hadoop), Guava (https://github.com/google/guava), Spring (https://github.com/spring-projects/spring-framework), Libgdx (https://github.com/libgdx/libgdx), Jenkins (https://github.com/jenkinsci/jenkins), Open Legislation (https://github.com/nysenate/OpenLegislation), Seleniun (https://github.com/nysenate/OpenLegislation), PrestoDB (https://github.com/prestodb/presto).



\section{References}


